---
title: "SLMPD Data Cleaning"
author: "Carter Hanford"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This is the R notebook for accessing data from the St. Louis Metropolitan Police Department.

```{r}
# load dependencies
library(compstatr)
library(dplyr)
library(here)
library(tidyr)
library(readr)
```

## Load Data
First we'll load the data from the **compstatr** package.

```{r}
# create index
i = cs_create_index()

# get years 
may10 = cs_get_data(year = 2010, month = "May", index = i)
may11 = cs_get_data(year = 2011, month = "May", index = i)
may12 = cs_get_data(year = 2012, month = "May", index = i)
may13 = cs_get_data(year = 2013, month = "May", index = i)
may14 = cs_get_data(year = 2014, month = "May", index = i)
may15 = cs_get_data(year = 2015, month = "May", index = i)
may16 = cs_get_data(year = 2016, month = "May", index = i)
may17 = cs_get_data(year = 2017, month = "May", index = i)
may18 = cs_get_data(year = 2018, month = "May", index = i)
```

Next let's select each year and features we want and combine everything into one dataset:

```{r}
# select variables
may10 %>%
  select(month_reportedto_mshp, crime, district, neighborhood, description, ileads_address, ileads_street) %>%
  unite(address, c(ileads_address, ileads_street), sep = " ") -> may10

may11 %>%
  select(month_reportedto_mshp, crime, district, neighborhood, description, ileads_address, ileads_street) %>%
  unite(address, c(ileads_address, ileads_street), sep = " ") -> may11

may12 %>%
  select(month_reportedto_mshp, crime, district, neighborhood, description, ileads_address, ileads_street) %>%
  unite(address, c(ileads_address, ileads_street), sep = " ") -> may12

may13 %>%
  select(month_reportedto_mshp, crime, district, neighborhood, description, ileads_address, ileads_street) %>%
  unite(address, c(ileads_address, ileads_street), sep = " ") -> may13

may14 %>%
  select(coded_month, crime, district, neighborhood, description, ileads_address, ileads_street) %>%
  rename(month_reportedto_mshp = coded_month) %>%
  unite(address, c(ileads_address, ileads_street), sep = " ") -> may14

may15 %>%
  select(coded_month, crime, district, neighborhood, description, ileads_address, ileads_street) %>%
  rename(month_reportedto_mshp = coded_month) %>%
  unite(address, c(ileads_address, ileads_street), sep = " ") -> may15

may16 %>%
  select(coded_month, crime, district, neighborhood, description, ileads_address, ileads_street) %>%
  rename(month_reportedto_mshp = coded_month) %>%
  unite(address, c(ileads_address, ileads_street), sep = " ") -> may16

may17 %>%
  select(coded_month, crime, district, neighborhood, description, ileads_address, ileads_street) %>%
  rename(month_reportedto_mshp = coded_month) %>%
  unite(address, c(ileads_address, ileads_street), sep = " ") -> may17

may18 %>%
  select(coded_month, crime, district, neighborhood, description, ileads_address, ileads_street) %>%
  rename(month_reportedto_mshp = coded_month) %>%
  unite(address, c(ileads_address, ileads_street), sep = " ") -> may18


# combine all dataframes
stlcrime <- rbind(may10, may11, may12,
                 may13, may14, may15,
                 may16, may17, may18)
```

Finally we'll filter for non-dangerous crimes only and then perform some organization:

```{r}
# filter for non-dangerous crimes
stlcrime %>%
  filter(description == "LOITERING-BEGGING" | description == "PUBLIC ORDER-OTHR UNSPC PBLC ORDER VIOLATION" | description == "LIQUOR LAWS-OTHR UNSPEC LIQ OFFNSE" | description == "DISORDERLY CONDUCT-PEACE DSTRB/INDIVIDUAL" | description == "DRUGS-POSSESSION/MARIJUANA" | description == "PUBLIC ORDER-TRESPASSING") -> stlcrime

# organize
stlcrime %>%
  rename(year = month_reportedto_mshp) %>%
  select(year, crime, district, neighborhood, address, description) -> stlcrime
```

And we'll write the cleaned STLMPD data to a csv:

```{r}
write_csv(stlcrime, here("data", "stlcrime.csv"))
```


















